---
title: "Geospatial Visualization with R"
subtitle: "From Spatial Data to Visual Insights"
author: "Erdem Özkul"
institute: "Hochschule Fresenius University of Applied Sciences"
date: today

format:
  revealjs:
    theme: simple
    slide-number: true
    transition: fade
    incremental: false
    width: 1050
    height: 700
  pdf:
    toc: false
    number-sections: false

execute:
  echo: false
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(sf)
library(ggplot2)
library(giscoR)
library(leaflet)
library(scales)
```

```{r}
#| include: false

germany <- gisco_get_countries(
  country = "DE",
  resolution = "20",
  spatialtype = "RG"
)

states <- gisco_get_nuts(
  country = "DE",
  nuts_level = 1,
  year = "2021",
  resolution = "20"
)

cities <- gisco_get_urban_audit(
  country = "DE",
  spatialtype = "LB",
  year = "2021"
)

states_density <- states %>%
  st_transform(3035) %>%
  mutate(
    area_km2 = as.numeric(st_area(.)) / 1000000,
    pop_density = case_when(
      NAME_LATN == "Berlin" ~ 4200,
      NAME_LATN == "Hamburg" ~ 2500,
      NAME_LATN == "Bremen" ~ 1600,
      NAME_LATN == "Nordrhein-Westfalen" ~ 530,
      NAME_LATN == "Saarland" ~ 380,
      NAME_LATN == "Hessen" ~ 300,
      NAME_LATN == "Baden-Württemberg" ~ 315,
      NAME_LATN == "Bayern" ~ 190,
      NAME_LATN == "Sachsen" ~ 220,
      NAME_LATN == "Rheinland-Pfalz" ~ 210,
      NAME_LATN == "Schleswig-Holstein" ~ 185,
      NAME_LATN == "Niedersachsen" ~ 170,
      NAME_LATN == "Thüringen" ~ 130,
      NAME_LATN == "Brandenburg" ~ 85,
      NAME_LATN == "Sachsen-Anhalt" ~ 105,
      NAME_LATN == "Mecklenburg-Vorpommern" ~ 70,
      TRUE ~ 150
    )
  ) %>%
  st_transform(4326)

cities_xy <- cities %>%
  st_transform(4326) %>%
  mutate(
    lon = st_coordinates(.)[,1],
    lat = st_coordinates(.)[,2]
  )
```

## Geospatial Visualization with R

### From Spatial Data to Visual Insights

<br>

**Data → Map → Insight → Decision**

---

## Why Location Matters

::: {.columns}

::: {.column width="33%"}

### Urban Planning

Land use  
Accessibility  
Infrastructure

:::

::: {.column width="33%"}

### Transportation

Traffic analysis  
Mobility patterns  
Route optimization

:::

::: {.column width="33%"}

### Real Estate

Housing prices  
Location value  
Market analysis

:::

:::

<br>

::: {.columns}

::: {.column width="50%"}

### Environment

Climate analysis  
Flood risk  
Green spaces

:::

::: {.column width="50%"}

### Business Intelligence

Site selection  
Customer distribution  
Market potential

:::

:::

---

## Spatial vs Geospatial Data

| Feature | Spatial Data | Geospatial Data |
|---|---|---|
| Reference | Any space | Earth |
| Coordinates | Optional | Required |
| Example | CAD drawing | GPS data |
| Use | Design / modeling | Mapping / analysis |

---

## Types of Geospatial Data

```{r}
par(mfrow = c(2,2), mar = c(1,1,3,1))

points <- st_as_sf(
  data.frame(x = c(1,2,3,2.5), y = c(1,2,1.5,3)),
  coords = c("x","y")
)

plot(st_geometry(points), pch = 16, cex = 2.5, main = "Point")

line <- st_sfc(
  st_linestring(matrix(c(1,1, 2,3, 3,2, 4,4), byrow = TRUE, ncol = 2))
)

plot(line, lwd = 4, main = "Line")

poly <- st_sfc(
  st_polygon(list(matrix(c(1,1, 4,1, 4,4, 1,4, 1,1), byrow = TRUE, ncol = 2)))
)

plot(poly, col = "grey85", border = "black", lwd = 2, main = "Polygon")

r <- matrix(runif(100), nrow = 10)

image(r, axes = FALSE, main = "Raster")
```

---

## Typical Workflow in R

<br><br>

# Import → Clean → Transform CRS → Visualize → Analyze

<br><br>

```{r}
#| eval: false
st_read()
st_transform()
ggplot()
```

---

## Main R Packages

| Package | Purpose |
|---|---|
| `sf` | Vector spatial data |
| `terra` | Raster data |
| `ggplot2` | Static maps |
| `leaflet` | Interactive maps |
| `osmdata` | OpenStreetMap data |
| `tmap` | Thematic mapping |

---

## Point Map

### German Cities

```{r}
ggplot() +
  geom_sf(data = germany, fill = "grey96", color = "grey50", linewidth = 0.3) +
  geom_sf(data = cities, size = 1.3, alpha = 0.75) +
  theme_void()
```

---

## Choropleth Map

### Population Density by State

```{r}
ggplot(states_density) +
  geom_sf(aes(fill = pop_density), color = "white", linewidth = 0.25) +
  scale_fill_viridis_c(
    option = "magma",
    labels = comma,
    name = "People / km²"
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )
```

---

## Heatmap

### Density of Urban Centres

```{r}
ggplot() +
  geom_sf(data = germany, fill = "grey96", color = "grey60", linewidth = 0.3) +
  stat_density_2d(
    data = cities_xy,
    aes(
      x = lon,
      y = lat,
      fill = after_stat(level),
      alpha = after_stat(level)
    ),
    geom = "polygon",
    bins = 14
  ) +
  scale_alpha(range = c(0.12, 0.8), guide = "none") +
  scale_fill_viridis_c(option = "inferno", guide = "none") +
  coord_sf(
    xlim = c(5.5, 15.5),
    ylim = c(47, 55.5),
    expand = FALSE
  ) +
  theme_void()
```

---

## Interactive Mapping

::: {.content-visible when-format="html"}

```{r}
leaflet(cities) %>%
  addTiles() %>%
  addCircleMarkers(
    radius = 4,
    stroke = FALSE,
    fillOpacity = 0.7
  )
```

:::

::: {.content-visible when-format="pdf"}

```{r}
ggplot() +
  geom_sf(data = germany, fill = "grey96", color = "grey50", linewidth = 0.3) +
  geom_sf(data = cities, size = 1.3, alpha = 0.75) +
  theme_void()
```

:::

---

## What Each Map Type Shows

| Map Type | Best For |
|---|---|
| Point Map | Exact locations |
| Choropleth Map | Regional comparison |
| Heatmap | Spatial density |
| Interactive Map | User exploration |

---

## Why It Matters

| From | To |
|---|---|
| Raw coordinates | Readable maps |
| Complex tables | Spatial comparison |
| Hidden patterns | Decision support |

---

## Conclusion

<br>

# Data → Maps → Insights → Decisions

<br>

Geospatial visualization with R helps analysts transform location-based data into clear and reproducible visual insights.

---

## Thank You

Questions?